---
title: "Spatial Randomforest"
author: "Christopher"
date: '2021-03-26'
slug: spatial-randomforest
categories: []
tags:
- Machine learning
- Meuse
- Soil Science
subtitle: ''
summary: This is a small tutorial to run randomForest on the Meuse Dataset
authors: []
lastmod: '2021-03-26T09:14:41-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to Run a RF Model

Although Random forest is a very powerful tool, it can be run fairly simply in R.
This tutorial can be followed using this [data](https://drive.google.com/file/d/1G6C2WhYcouP5GAUE27WwQtXlsjMgdf22/view?usp=sharing) 

```{r message=FALSE, warning=FALSE}
#Lets read our libraries in first then load our shape file
library(rgdal)
library(raster)
library(sp)
library(randomForest)
library(tidyverse)

#We are going to extract points from a raster to a shapefile

WorkWD <- "C:/Users/User/Documents/R/Projects/R_Intro"

Training <- readOGR(dsn = WorkWD, layer = "TrainingData")

plot(Training)
```

## Coordinate Reference System and Spatial Projection

R is very picky about how we project our data, along with the geographical projection

```{r}
#Apply a consitent CRS- Coordinate Reference System
proj4string(Training) <- CRS("+init=epsg:28992")

#The geographic projection has to be changed as well
Training  <- spTransform(Training , CRS("+init=epsg:28992"))

```

## Raster Data

Lets load in our raster data and see what it looks like.

```{r}
#Load in the DEM and extract elevation values

#LOADINING THE RASTERS
DEM <- raster( file.path(WorkWD, "Elevationlow.tif"))
River <- raster(file.path(WorkWD, "Distance.tif"))

#lets plot the DEM to see what we have

plot(DEM)

#Apply a consitent CRS
proj4string(DEM)<- CRS("+init=epsg:28992")
proj4string(River)<- CRS("+init=epsg:28992")

#Crop the River Raster to fit the DEM
#
River <- crop(River,DEM)
```

## Raster conversion and extract:

We can compute the slope and aspect as well The final function is the extract function, similar to the "extract multi values" tool in ArcGIS

```{r}
#We are also creating a slope and aspect layer
Slope <- terrain(DEM, opt="slope", unit="degrees", neighbors=8)
Aspect <- terrain(DEM, opt="aspect", unit="degrees", neighbors=8)


#Extract Elevation Values
#raster::extract is saying to use the extract function directly from our raster package
Training$DEM <- raster::extract(DEM, Training, method = "simple")
Training$Slope <- raster::extract(Slope, Training, method = "simple")
Training$Aspect <- raster::extract(Aspect, Training, method = "simple")
Training$River <- raster::extract(River, Training, method = "simple")

#Lets check to see the summary statistics of our newly extracted data
summary(Training)

```

## Spatial objects to Data frames

We need to convert our spatial object into a `data.frame` if we want to conduct more analysis on our data

```{r}
#Spatial objects hold a dataframe, but they are treated like a spatial object, we need to convert the object to a dataframe in order to do additional analysis
Training.DF <- as.data.frame(Training)

#lets see the new structure of that DF
str(Training.DF)

#drop the NA values so that we can safely run our RF

Training.DF <- Training.DF %>%  
                drop_na()

```

## Running Random Forest

In order to run random forest, we need to use our `data.frame` and decide which are the independant and dependent variables.

```{r}
set.seed(95)

MeuseRF<- randomForest(lead ~ DEM+Aspect+Slope+River, data=Training.DF, importance=TRUE, proximity=FALSE, varImpPlot = TRUE, varUsed = TRUE, TYPE=regression, ntree=1000)

#To see the r2 Value
MeuseRF

```

## Plotting the RF outputs

The plots that we need are the var

```{r}

#to see the variable importance plots
varImpPlot(MeuseRF, main="All Variables")

partialPlot(MeuseRF, Training.DF, River, , main = "PPD River")
partialPlot(MeuseRF, Training.DF, Slope, , main = "PPD Slope")
partialPlot(MeuseRF, Training.DF, DEM, , main = "PPD Elevation")


#to predict the RF to an output, we have to create a raster stack of our data so that it cant determine the optimal placement of values

```

## Predicting the output

to predict the RF to an output, we have to create a raster stack of our data so that it can determine the optimal placement of values

```{r}

#Creating the Raster Stack
Stack.Meuse <-stack(DEM,Slope,Aspect,River)

#The names of the Raster Stack and the Training Dataframe were not lining up
names(Stack.Meuse) <- c("DEM","Slope","Aspect","River")

Lead <- predict(Stack.Meuse, MeuseRF, filename = "lead.tif", fun = predict, se.fit=TRUE, overwrite=TRUE)

plot(Lead, main= "Lead for the Meuse Data")


```
# Some final interpretations 
The final output has a band which resembles the distance from river layer. This would bring into question the accuracy of the lead layer (kriging would probably be a more appropriate usage here). Test the code on your own data and see how it looks.